---
title: "Project_mini_GZL"
author: "Gabriela Cromhout"
date: "2025-02-13"
output:
  slidy_presentation: default
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```




```{r}
library(tidyverse)
library(dplyr)
library(dslabs)
maternal_data <- read.table(file = 'Maternal Health Risk Data Set.csv', header = TRUE, sep = ",",
                        stringsAsFactors = TRUE)
```
```{r}
head(maternal_data)
```
```{r}
maternal_data %>% 
  group_by(RiskLevel) %>%
  summarise(avg = mean(Age))%>%
  arrange(desc(avg))

```




#Q1ii
```{r}
table_dias <- select(maternal_data, RiskLevel, DiastolicBP)
```
```{r}
dias_table <- select(maternal_data, RiskLevel, DiastolicBP)
```

```{r}
dias_table %>% group_by(RiskLevel) %>%summarise(avg=median(DiastolicBP)) %>%   arrange(desc(avg))
```
#Q2 New variable - systolic rate
```{r}
maternal_data_1 <- mutate(maternal_data, systolic_rate = SystolicBP/Age)
head(maternal_data_1)
```

#Q3 - filter for <40

#filter then SBP by risk level

```{r}
filter(maternal_data_1, Age<=40) %>% group_by(RiskLevel) %>% summarise(avg=mean(SystolicBP)) %>% arrange(desc(avg))
```

#Q4 - histo & boxplot Age
```{r}
hist(maternal_data$Age, 
     col = "lightblue",  # Fill color of bars
     border = "black",   # Border color of bars
     main = "Age Distribution",
     xlab = "Age",
     ylab = "Frequency")
```
#Majority of the data is in the 10-35 age group
```{r}
boxplot(maternal_data$Age)
```
#histo & boxplot SBP
```{r}
hist(maternal_data$SystolicBP,
      col = "lightblue",  # Fill color of bars
     border = "black",   # Border color of bars
     main = "Histogram of Systolic BP",
     xlab = "Systolic",
     ylab = "Frequency")
```
```{r}
boxplot(maternal_data$SystolicBP)
```
```{r}
fivenum(maternal_data$SystolicBP)
```
#histo & boxplot DBP
```{r}
hist(maternal_data$DiastolicBP, 
      col = "lightblue",  # Fill color of bars
     border = "black",   # Border color of bars
     main = "Diastolic BP Distribution",
     xlab = "Diastolic BP",
     ylab = "Frequency")
```
```{r}
boxplot(maternal_data$DiastolicBP)
```
```{r}
fivenum(maternal_data$DiastolicBP)
```
# histo & boxplot BS
```{r}
hist(maternal_data$BS,
      col = "lightblue",  # Fill color of bars
     border = "black",   # Border color of bars
     main = "Blood Sugar Distribution",
     xlab = "Blood sugar",
     ylab = "Frequency")
```
```{r}
boxplot(maternal_data$BS)
```
```{r}
fivenum(maternal_data$BS)
```
#histo & boxplot body temp
```{r}
hist(maternal_data$BodyTemp,
      col = "lightblue",  # Fill color of bars
     border = "black",   # Border color of bars
     main = "Body Temp Distribution",
     xlab = "Body Temp",
     ylab = "Frequency")
```
```{r}
boxplot(maternal_data$BodyTemp)
```
```{r}
fivenum(maternal_data$BodyTemp)
```

#histo & boxplot HR
```{r}
hist(maternal_data$HeartRate,
      col = "lightblue",  # Fill color of bars
     border = "black",   # Border color of bars
     main = "Heart Rate Distribution",
     xlab = "Heart Rate",
     ylab = "Frequency")
```
```{r}
boxplot(maternal_data$HeartRate)
```

#correct grouping of 5nums acc to risk level
#HR
```{r}
tapply(maternal_data$HeartRate, maternal_data$RiskLevel, summary)
```
#Age

```{r}
tapply(maternal_data$Age, maternal_data$RiskLevel, summary)
```
#SBP
```{r}
tapply(maternal_data$SystolicBP, maternal_data$RiskLevel, summary)
```
#DBP
```{r}
tapply(maternal_data$DiastolicBP, maternal_data$RiskLevel, summary)
```
#BodyTemp
```{r}
tapply(maternal_data$BodyTemp, maternal_data$RiskLevel, summary)
```
#BS
```{r}
tapply(maternal_data$BS, maternal_data$RiskLevel, summary)
```
#Age vs HR scatterplot
```{r, eval=FALSE}
with(maternal_data, plot(HeartRate, Age))

```
#SBP vs HR scatterplot
```{r, eval=FALSE}
with(maternal_data, plot( SystolicBP, HeartRate))
```
#DBP vs HR
```{r}
with(maternal_data, plot( DiastolicBP, HeartRate))
```
#Body temp vs HR
```{r}
with(maternal_data, plot(BodyTemp, HeartRate))
```
```{r}
# Reshape the data to long format
data_long <- maternal_data %>%
  select(Age, SystolicBP, DiastolicBP, BS, BodyTemp, HeartRate, RiskLevel) %>%
  pivot_longer(cols = c(Age, SystolicBP, DiastolicBP, BS, BodyTemp, HeartRate),
               names_to = "Measurement", values_to = "Value")

# Corrected histogram plot (x-axis should be continuous)
ggplot(data_long, aes(x = Value, fill = RiskLevel)) +
  geom_histogram(binwidth = 5, alpha = 0.6, position = "identity") +
  facet_wrap(~ Measurement, scales = "free_x") +
  labs(title = "Histograms of Measurements", 
       x = "Value", y = "Count") +
  theme_minimal()
  

```
```{r}

library(corrplot)

# Compute correlation matrix
cor_matrix <- cor(maternal_data[, c("Age", "DiastolicBP", "SystolicBP")])

# Plot the correlation matrix
corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8)


```
```{r}
vars_to_plot <- c("Age", "SystolicBP", "DiastolicBP", "BodyTemp") 

for (var in vars_to_plot) {ggplot(maternal_data, aes(x = .data[[var]], y = HeartRate)) + geom_point(alpha = 0.6, color = "red") + geom_smooth(method = "lm", se = FALSE, color = "blue") +labs(title = paste("Heart Rate vs", var), x = var, y = "Heart Rate") + theme_minimal() -> p 
  print(p)}

```


#Linear model HR vs BT
```{r}
lm_BT = lm(formula = HeartRate ~ BodyTemp, data = maternal_data)
summary(lm_BT)
```
#Linear model HR vs DBP
```{r}
lm_DBP = lm(formula = HeartRate ~ DiastolicBP, data = maternal_data)
summary(lm_DBP)
```
#Linear model HR vs SBP
```{r}
lm_SBP = lm(formula = HeartRate ~ SystolicBP, data = maternal_data)
summary(lm_SBP)
```
#Linear model HR vs Age
```{r}
lm_age = lm(formula = HeartRate ~ Age, data = maternal_data)
summary(lm_age)
```
#Linear model BS vs HR
```{r}
lm_BS = lm(formula = HeartRate ~ BS, data = maternal_data)
summary(lm_BS)
```
#Multiple lm  HR
```{r}
multi_lm = lm(formula = HeartRate ~ .,
               data = maternal_data)
summary(multi_lm)

```
#reworked
```{r}
multi_lm_1 = lm(formula = HeartRate ~ DiastolicBP + Age + BodyTemp + SystolicBP,
               data = maternal_data)
summary(multi_lm_1)

```
#step-wise - both directions
```{r}
intercept_only <- lm(HeartRate ~ 1, data=maternal_data) 
```

```{r}
all <- lm(HeartRate ~ SystolicBP + DiastolicBP + BodyTemp + Age, data=maternal_data)
```

```{r}
both <- step(intercept_only, direction = 'both', scope = formula(all), trace=0)
```

```{r}
both$anova
```
#both coefficient - explains relationship
```{r} 
both$coefficients
```