---
title: "Project_mini_GZL"
author: "Gabriela Cromhout"
date: "2025-02-13"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

```{r, eval=FALSE}
getwd()
```

```{r}
library(tidyverse)
library(dplyr)
library(dslabs)
maternal_data <- read.table(file = 'Maternal Health Risk Data Set.csv', header = TRUE, sep = ",",
                        stringsAsFactors = TRUE)
```
```{r}
head(maternal_data)
```
#Q1ii
```{r}
table_dias <- select(maternal_data, RiskLevel, DiastolicBP)
```
```{r}
dias_table <- select(maternal_data, RiskLevel, DiastolicBP)
```
```{r}
dias_table
```
```{r}
dias_table %>% group_by(RiskLevel) %>%summarise(avg=median(DiastolicBP)) %>% arrange(avg)
```
#Q2 New variable - systolic rate
```{r}
maternal_data_1 <- mutate(maternal_data, systolic_rate = SystolicBP/Age)
```
```{r}
maternal_data_1
```
#Q3 - filter for <40
```{r}
filter(maternal_data_1, Age<=40)
```
#filter then SBP by risk level

```{r}
filter(maternal_data_1, Age<=40) %>% group_by(RiskLevel) %>% summarise(avg=mean(SystolicBP)) %>% arrange(avg)
```
```{r}
head(maternal_data)
```
#Q4 - histo & boxplot Age
```{r}
hist(maternal_data$Age)
```
#Majority of the data is in the 10-35 age group
```{r}
boxplot(maternal_data$Age)
```
#histo & boxplot SBP
```{r}
hist(maternal_data$SystolicBP)
```
```{r}
boxplot(maternal_data$SystolicBP)
```
```{r}
fivenum(maternal_data$SystolicBP)
```
#histo & boxplot DBP
```{r}
hist(maternal_data$DiastolicBP)
```
```{r}
boxplot(maternal_data$DiastolicBP)
```
```{r}
fivenum(maternal_data$DiastolicBP)
```
# histo & boxplot BS
```{r}
hist(maternal_data$BS)
```
```{r}
boxplot(maternal_data$BS)
```
```{r}
fivenum(maternal_data$BS)
```
#histo & boxplot body temp
```{r}
hist(maternal_data$BodyTemp)
```
```{r}
boxplot(maternal_data$BodyTemp)
```
```{r}
fivenum(maternal_data$BodyTemp)
```

#histo & boxplot HR
```{r}
hist(maternal_data$HeartRate)
```
```{r}
boxplot(maternal_data$HeartRate)
```
```{r}
summary_risk <- maternal_data  %>% group_by(RiskLevel) %>% summarise(fivenum(summary_risk)) %>% print(summary_risk)
```
```{r}
summarise(maternal_data, RiskLevel, HeartRate)
```
```{r}
group_by(RiskLevel) %>% summarise(5num = fivenum(HeartRate))
```
#correct grouping of 5nums acc to risk level
#HR
```{r}
tapply(maternal_data$HeartRate, maternal_data$RiskLevel, summary)
```
#Age

```{r}
tapply(maternal_data$Age, maternal_data$RiskLevel, summary)
```
#SBP
```{r}
tapply(maternal_data$SystolicBP, maternal_data$RiskLevel, summary)
```
#DBP
```{r}
tapply(maternal_data$DiastolicBP, maternal_data$RiskLevel, summary)
```
#BodyTemp
```{r}
tapply(maternal_data$BodyTemp, maternal_data$RiskLevel, summary)
```
#BS
```{r}
tapply(maternal_data$BS, maternal_data$RiskLevel, summary)
```
#Age vs HR scatterplot
```{r, eval=FALSE}
with(maternal_data, plot(HeartRate, Age))
```
#SBP vs HR scatterplot
```{r, eval=FALSE}
with(maternal_data, plot( SystolicBP, HeartRate))
```
#DBP vs HR
```{r}
with(maternal_data, plot( DiastolicBP, HeartRate))
```
#Body temp vs HR
```{r}
with(maternal_data, plot(BodyTemp, HeartRate))
```
#Linear model HR vs BT
```{r}
lm_BT = lm(formula = HeartRate ~ BodyTemp, data = maternal_data)
summary(lm_BT)
```
#Linear model HR vs DBP
```{r}
lm_DBP = lm(formula = HeartRate ~ DiastolicBP, data = maternal_data)
summary(lm_DBP)
```
#Linear model HR vs SBP
```{r}
lm_SBP = lm(formula = HeartRate ~ SystolicBP, data = maternal_data)
summary(lm_SBP)
```
#Linear model HR vs Age
```{r}
lm_age = lm(formula = HeartRate ~ Age, data = maternal_data)
summary(lm_age)
```
#Linear model BS vs HR
```{r}
lm_BS = lm(formula = HeartRate ~ BS, data = maternal_data)
summary(lm_BS)
```
#Multiple lm  HR
```{r}
multi_lm = lm(formula = HeartRate ~ .,
               data = maternal_data)
summary(multi_lm)

```
#reworked
```{r}
multi_lm_1 = lm(formula = HeartRate ~ DiastolicBP + Age + BodyTemp + SystolicBP,
               data = maternal_data)
summary(multi_lm_1)

```
#step-wise - both directions
```{r}
intercept_only <- lm(HeartRate ~ 1, data=maternal_data) 
```

```{r}
all <- lm(HeartRate ~ SystolicBP + DiastolicBP + BodyTemp + Age, data=maternal_data)
```

```{r}
both <- step(intercept_only, direction = 'both', scope = formula(all), trace=0)
```

```{r}
both$anova
```
#both coefficient - explains relationship
```{r} 
both$coefficients
```